---
html_document:
  toc: yes
author: "Unstructured Data Analytics, Fall 2023"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Analyzing Text: Preliminaries'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Well well well, What do we have here? 
*First explorations of a new dataset*

```{r echo=TRUE, message=FALSE, warning=FALSE}
#load initial packages
library(tidyverse)
library(stringr)
```

  
## Let's read in some data and check it out

```{r}
#read it in
sports <- read_csv("???/UDA Fall 2023 Stories.csv")
```
How can we explore this dataset to find out what it offers us?
```{r}
#explore!

```


## "Bag of words" methods

*How do we find the words?*
*What issues might we run into?*

### Dividing and Counting
Can you count how many words there are in one Story? In all the stories?
Can you tell me how many times a particular word shows up in one Story? In all the stories? 
```{r}
#let's count

```
## Is there an easier way?
Yes, of course, this is R.
```{r}
#load package

```
Divide and prep
```{r}
#use package

```
What changes were made from the starting text to the version? 
```{r}
#write code to explore as needed


```


## Counts are surprisingly useful

### Which words are most common overall? 
### Which words are most common in each story? 

Can you create a top-five list of words for the whole file? 
Can you create a top-five list of words for each document?
*Store these as new objects called "top_by_file" and "top_by_doc"*

```{r}
#everyone loves a "top" list

```

## Can we get a better list of common words?
Yes, of course. 

```{r}
#load the data

```
How many lexicons are there? 
How many words are there in each lexicon?
*Do you want more dictionary options? Try the "stopwords" package.  It has additional stop words dictionaries, including support for foreign languages.*
```{r}
#explore lexicons
```

What happens to our counts if we remove the stopwords using the SMART lexicon?
```{r}
#remove SMART words

```
Can we see which stopwords were removed?  
(Are there any we might want to keep?)
```{r}
#review what we lost

```

