---
html_document:
  toc: yes
author: "Unstructured Data Analytics, Fall 2024"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Analyzing Text: Sentiment Analysis, Part 2'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Sentiment: Graphing Polarity


#### Prep code, from before -- run!
```{r}
## copy in from Day 06B completed
```



### How might we graph polarity?
"up-down" bar chart
```{r}

```

### Graphing if you have many neutral values?
scatterplot
```{r}

```


# Sentiment with emotion

#### Which emotions are present in each story?
#### Which emotion is the most prevalent? What is the mix like? 

Which lexicon do we need?
```{r}
#which emotion is the most prevalent in each story?

```


```{r}
#what is the mix of emotions like in each story?

```


#### Let's visualize the mix of emotions within a document
```{r}
library(RColorBrewer)  #for more color options

dat_words %>%
  inner_join(nrc, by = "word") %>%
  count(id, feeling, sentiment) %>%
  filter(!sentiment %in% c("negative", "positive")) %>%
  ggplot(
    aes(x = id, y = n, fill = sentiment)
  ) +
  geom_col(position = "fill") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip() +
  theme_minimal() +
  ggtitle("Sentiment range for each story")

```



# Sentiment: Valence

#### How can we start to incorporate context to help with valence?
#### Can we view sentiment at a higher level, beyond single words, before we aggregate?

Yes, of course we can!  

This capability lets us account (to some extent) for valence shifters.

### Implementing higher-level sentiment
The tidytext package is set up for word-level sentiment only.  That means we have to shift to the `sentimentr` package.  
  
Let's find new sentiment values for each story, taking valence into consideration
```{r}
library(sentimentr)


```

How can we easily check for example sentences or review the scoring more visually?
```{r}

```

# Challenge
Instead of going story-by-story, can we look at these stories by different groups? 
What groups might makes sense? Set it up, and redo sentiment analysis to get more general insights (if any!)

```{r}

```

