---
html_document:
  toc: yes
author: "Unstructured Data Analytics, Fall 2024"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Analyzing Text: Sentiment Analysis'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# How can we detect if documents contain positive/negative feeling or specific emotion?


## First Step: Analyze Polarity with Lexicons
```{r echo=TRUE, message=FALSE, warning=FALSE}
# load initial packages
library(tidyverse)
library(tidytext)
```

First step: getting lexicons  
*For some, you have to download them: switch to the console to approve download*
```{r}
# this one for free
bing <- get_sentiments(lexicon = "bing")

# also load two more!

```

  
### *Question: How do these sentiment lexicons vary in size and structure?*   
Let's take a look
```{r}


```

  
## Let's keep our first focus on polarity

### Question: If we want to use the afin data to score our stories, what general process should we use?

### Question: If we want to use the bing data to score our stories, what is different?


## Implementation
First, we have to get our text data ready

```{r}
library(googledrive)
library(googlesheets4)


#read in
drive_auth()
gs4_auth(token = drive_token())
ss <- drive_get("UDA 2024 South Bend")
dat <- read_sheet(ss)

#explore
glimpse(dat)

```

 
Cleanup, as needed.  Prep appropriately for our next step
```{r}


```

Combine words with sentiment: afinn
```{r}

```

### Question: What is the polarity of each story?

```{r}

```
### What do the polarity numbers represent? 

```{r}

```


### What results do we get using bing?
```{r}

```


## How might we graph polarity?
"up-down" bar chart
```{r}

```

## Graphing if you have many neutral values?
scatterplot
```{r}

```



## Which emotions are present in each story?
### Which emotion is the most prevalent? What is the mix like? 

Which lexicon do we need?
```{r}
#which emotion is the most prevalent in each story?

```


```{r}
#what is the mix of emotions like in each story?

```


### Let's visualize the mix of emotions within a document
```{r}
library(RColorBrewer)  #for more color options

dat_words %>%
  inner_join(nrc, by = "word") %>%
  count(id, feeling, sentiment) %>%
  filter(!sentiment %in% c("negative", "positive")) %>%
  ggplot(
    aes(x = id, y = n, fill = sentiment)
  ) +
  geom_col(position = "fill") +
  scale_fill_brewer(palette = "Set1") +
  coord_flip() +
  theme_minimal() +
  ggtitle("Sentiment range for each story")

```

