---
title: 'Wrangling Review with R: Data manipulation'
output:
  html_document:
    df_print: paged
---

## Instructions
Add code below to perform the requested actions.  
Submit your final result as a knitted HTML file to Canvas.  

I am hoping that Q1-Q4 are familiar and will take no more than 15-20 minutes
I expect that Q5+ may have some things you don't know.  Take a look and try them, but don't spend a lot of time if you're stuck. You can adjust the code as needed to get a (successful) knitted file.

Free startup
```{r startup, include = F}
library(dplyr)
library(tidyr)
glimpse(starwars)
```


### Q1
Limit the starwars dataset to only include the columns  
*name, height, mass, hair_color, eye_color, skin_color, and species*  
and assign it to the object sw  
Hint for this and following: https://r4ds.had.co.nz/transform.html
```{r Q1}

sw <- starwars #your code here

#show the output
colnames(sw)
```

### Q2
Using your new dataset, how many individuals have each hair color?
```{r Q2}
#your code here

```


### Q3
Limit your new dataset *sw* further to **only** include characters who are taller than 181cm
```{r Q3}
sw_taller <- sw #your code here

#show some output
min(sw_taller$height, na.rm = TRUE)
```


### Q4
Create a new column for the base starwars data that converts cm to inches, called "height_inch"  
The value should be height divided by 2.54 
```{r Q4}
starwars #your code here

#show some results
paste("Characters who are", starwars$height[1], "in cm are", starwars$height_inch[1], "in inches")
```


## This is where it might start to get harder: we'll review these things in class as needed

### Q5
Order your revised dataset alphabetically by name, and show the first 10
```{r Q5}
sw_names <- sw  #your code here

#show some output (adjust/add to this code if necessary)
head(sw_names, 10)
#or
sw_names[1:10,]
```



### Q6
For each species as a group, calculate and show the *average height in inches* and the *average mass*.  
```{r Q6}
sw_metrics <- #your code here

#show some results 
sw_metrics
```


### Q7 Make the sw data more vertical/taller (pivot longer) 
...by combining the information from eye_color, hair_color, and skin_color  
into one column (called "descriptor" with values like "eye_color" and "hair_color")   
and a second column (called "color" with values like "blue" and "blond").  
Hint: https://r4ds.had.co.nz/tidy-data.html

*Honestly, I don’t know how to do this in Base R successfully. Tidyverse or bust!
```{r Q7}
sw_pivoted <- sw  #your code here

#show some results
sw_pivoted
```


