---
html_document:
  toc: yes
author: "Your Name Here"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Assignment #2 - Fall 2023 - Insights from Counting'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Preliminaries
Load basic packages and data 
```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
```

Bring in data and explore a little. Decide what you want to analyze. Set up a new column for high/medium/low engagement.
```{r}
dat <- read_csv("Your CSV file from Assignment #1.csv")
```



# Counting

### Find "most common words" for the corpus (all together or grouped) 
Preprocess and get counts/frequencies
```{r}

```

### Find "most common phrases" (all together or grouped)
Preprocess and get counts/frequencies
```{r}

```

### Find "most important words" for the engagement levels
Preprocess and get tf-idf values based on engagement
```{r}

```

# Final Recommendation -- Thoughts and Visual
*Find the flowers amid all the weeds.* 

```{r}

```

