---
html_document:
  toc: yes
author: "Your Name Here"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Assignment #4 - Fall 2023 - Finding Latent Topics'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Preliminaries
Load basic packages and data 
```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
```

### Bring in data.
```{r}
dat <- read_csv("some reddit data.csv")
```


### Set up the data into the correct format for topic modeling

```{r}

```

## Run a topic model & variations (set a seed)
```{r}

```

## Compare results from the models
What emerges? 
```{r}

```
### Which model is best? Why? Can you visualize something about it?

### What is a top document from each topic?  
```{r}

```

## Final Thoughts
What did you expect to find? What did you actually find?  What names or themes would you assign to each topic? How coherent are they? Are the "top documents" actually good representatives for your topics? Would they be useful for MarComm to highlight in an fan engagement campaign?
