---
html_document:
  toc: yes
author: "Your Name Here"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Assignment #4 - Fall 2023 - Finding Latent Topics'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Preliminaries
Load basic packages and data 
```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
```

### Bring in data.
```{r}
dat <- read_csv("Your CSV file from Assignment #1.csv")
```


### Set up the data into the correct format for topic modeling

```{r}

```

## Run a topic model & variations (set a seed)
```{r}

```

## Compare results from the models
What emerges? 
```{r}

```
### Which model is best? Why? Can you visualize something about it?


## Final Thoughts
What did you expect to find? What did you actually find?  What names or themes would you assign to each topic? How coherent are they? 
