---
html_document:
  toc: yes
author: "Your Name Here"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Assignment #3 - Fall 2023 - Insights from Sentiment'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Preliminaries
Load basic packages and data 
```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
```

Bring in data.
```{r}
dat <- read_csv("Your CSV file from Assignment #1.csv")
```



# Sentiment Basics

### Find polarity-related sentiment scores for your data
Get a single score for each post.  

```{r}

```


Create an overview of the sentiment in the corpus 

 * If you just looked at polarity (positive vs negative), show how many posts were positive vs negative overall.  
 * If you looked at the strength of the polarity (5 to -5), break your scores into 3 or more groups (for example: positive, neutral, negative), and show how many posts fall into each group. 
 
You can show this using an aggregated dataframe/table or a graph.
```{r}

```
 
   
### Find the emotions in your data
For each post, detect emotions.  Then aggregate the whole corpus together.
Show how many times each emotion appears through the whole corpus.  Is it balanced? Do some emotions dominate?  ("Show" method can be a dataframe/table or a graph)
```{r}

```

# Sentiment deeper dive

Select either polarity or emotions. Examine the relationship between sentiment and some other element of this dataset. The other element should be grouped or aggregated in some way, for example, all posts with the same number of comments, comments divided into 5 different groups, time by day/week/month/weekday, all posts together with top words, etc.


#### WRITE: Describe your goal (what question do you want to look at?)


```{r}
#data prep code


```

```{r}
#visualization code


```




# Final Thoughts
*This can be commentary only or commentary plus code/visual*  
*It does not need to be long.  Quality over quantity*

