---
html_document:
  toc: yes
author: "Unstructured Data Analytics, Fall 2023"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'UDA Challenge Lab: Processing & Visualizing Text'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

#Today's Main Question:  

## Preliminaries
Load basic packages
```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
```


# Start with the revised data: Prep it

You'll want to work in steps today

 1. Read in the revised data file as `dat`
 2. Tokenize by words and store the result as `dat_words`
 3. Create overall counts (all documents together) in `dat_counts`
 4. Create story-by-story counts in `story_counts`
 5. *Challenge:* Create sport-by-sport counts in `sport_counts`

We want 3-4 different dataframes so we can compare things below
```{r}
#read in data


#explore


#tokenize


#create count-based data frames


```



  
# Task 1: ggplot to visualize count data
Note: the behavior of your data below will depend a little bit on how you create counts. If you use `group_by`, your data will retain a grouped structure by default.  If you use `count`, your data will be treated like a standard dataframe instead of structurally grouped.  
  
## Plot 1

*Start with the full data: `dat_counts` and build up your code piece by piece*
  
 * Can you create a bar chart that shows words and their wordcounts?  
 * Can you limit your bar chart to show only the top (x) words? (Pick a value for X)
 * Can you customize the title?
 * *Challenge:* Can you flip your axes so the words are on the y axis and the counts are on the x axis?


```{r}
#ggplot bar chart #1

```

#### Question:  Are there any key subjects or ideas that pop out when you look at the most common words?  
    
  
## Plot 2

*Switch to the `story_counts` or `sport_counts` data* 
  
+Start with the same code as above; consider a smaller value of (X) for space and legibility
+*Challenge:* Do you see each sport separately? If not, can you add a `facet_wrap` to your plot? 
```{r}
#ggplot bar chart #2

```


#### Question: Pick 2-3 stories or sports and look at their keywords in particular.  Is there any kind of theme that comes out?  
Remember that our top words for Story 1 showed that family and time were key ideas, through top words like "brother" and "year".   
*This kind of analysis works better with a lot more data, but you get the idea*  



#### Question: Do the top words tend to help you identify the sports? Explain with an example. Why might this happen (or not happen)?
 


#### Question: Compare the top words in the small graphs to each other, or to the larger graph (with all the data together).  Comment on something interesting.  Literally anything intersting.  


# Task 2: Create a wordcloud to visualize your data

Yes, you can do this manually in ggplot by setting different random x and y values for each word and then replacing a scatterplot point with the word itself but... that's a lot of work.

Instead. use this new-to-you package
```{r echo = FALSE, eval = FALSE}
# you'll probably need to install this package
# this chunk of code will be ignored when you knit 
install.packages("wordcloud")
```

```{r}
#load new package
library(wordcloud)

#view help
help(wordcloud)
```

Take a look at the help file for the wordcloud function.  
  
  
You'll need to use a minimum of two arguments: 

1. the dataframe column that has your words for the "words" argument (use $ notation in the function)  
2. the dataframe column that has your counts for the "freq" argument (use $ notation in the function)  
  
After you get those two arguments working

3. Try adding an argument to adjust the color
4. Try adding an argument to reduce the total number of words


*Use the dat_counts data*

```{r}
#your wordcloud here
```

Explore some variation in the package. Try the same wordcloud, but omit the freq argument and see what happens
```{r}
#your no-freq wordcloud here
```

#### Explain the difference between the two wordclouds.  Is the difference discussed in the help page documentation? 







  
