---
html_document:
  toc: yes
author: "Unstructured Data Analytics, Fall 2023"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Some thoughts on dates'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
# load initial packages
library(tidyverse)
```


# How can we easily work with dates? 


## First Step: Check the datatype of your date
```{r}
# I am also going to invent some data

df <- data.frame(DateText = c("2016-07-28", "2016-07-29", "2016-07-30", "2016-07-31", "2016-08-01", "2016-08-02")) %>%
  mutate(DateDate = as.Date(DateText))

#look at the data -- pay attention to the column type under the column name
df

```

If it says "date", congratulations!  If it says character, we need to convert it. The lubridate package is really easy to use for this.

### lubridate to convert text to dates
```{r}
library(lubridate)

df %>% 
  mutate(DateTextFixed = ymd(DateText))

```
The ymd function reads the text format in as year-month-day. For other options, see your lubridate cheatsheet  


## With dates: now what?
You can use date columns like other variables for grouping your text data, graphing, etc, the same way you might use a column called "Sport" or "ID.  

## Can I group the dates themselves?  
Absolutely.  There are many functions that will pull out a single part of your date.
 
### Date parts
```{r}
#can you see what each function is doing to our dates?
df %>%
  mutate(date_monthno = month(DateDate),
         date_weekno = week(DateDate),
         date_days = wday(DateDate))
```
 * `month` extracts the month as a number (July = 7, August = 8)
 * `week` counts weeks from the January 1st (we have Week #30 and #31 in this data)
 * `wday` uses a number to express a day of the week (Sunday = 1)
 
You can use these extractions to group data across multiple years into months, days, weeks (All "Decembers" together, for example).  

### Date rounding
Maybe you want to group dates within a single period together, like within the same week or within the same month, but you want to keep years apart.  
Try the floor and ceiling functions
```{r}
df %>%
  mutate(week_down = floor_date(DateDate, unit = "week"),
         month_down = floor_date(DateDate, unit = "month"),
         month_up = ceiling_date(DateDate, unit = "month"))
         
```
Compare each of the values to the original DateDate to see how we are rounding. We can use these rounded dates to help group our data into specific weeks or months in way that still preserves the year. 

### But wait, there's more!
lubridate offers functions to calculate durations, do math with dates, and to convert numeric representations of datatimes into timestamped dates (that's what you have in your "timestamp" reddit data -- try putting that column into the lubridate `as_datetime()` function).  

Learn more about these options on your cheatsheet or by looking online.  Lubridate makes it easy!