---
html_document:
  toc: yes
author: "Your Name Here"
output:
  html_document:
    df_print: paged
always_allow_html: yes
title: 'Assignment #3 - Fall 2023 - Insights from Sentiment'
toc_float: yes
theme: spacelab
highlight: tango
df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Preliminaries
Load basic packages and data 
```{r echo = TRUE, warning = FALSE, message = FALSE}
library(tidyverse)
library(tidytext)
```

Bring in data.
```{r}

```



# Sentiment Basics

### Find polarity-related sentiment scores for your data
Get a single score for each item  

```{r}

```


### Create an overview of the sentiment in the corpus 
This process requires two different rounds of aggregation.

 * If you just looked at polarity (positive vs negative), show how many **items (posts, reviews, etc)** were positive vs negative overall.  
 * If you looked at the strength of the polarity (5 to -5), break your scores into 3 or more groups (for example: positive, neutral, negative), and show how many items fall into each group. [If all of your posts are positive, consider "less, more, very positive" as groups] 
 
You can show the final situation (counts of items in a category) using an aggregated dataframe/table or a graph.
```{r}

```
 
   
### Find the emotions in your data
For each item, detect emotions.  Then look at all of your data together.
Show how many times each emotion appears through the whole corpus (all together).  Is it balanced? Do some emotions dominate?  ("Show" method can be a dataframe/table or a graph)
```{r}

```

# Sentiment deeper dive

Select either polarity or emotions. Examine the relationship between that sentiment and some other element of this dataset. The other element should be grouped or aggregated in some way, for example, by time or category or person or another relevant variable or really *anything* that seems interesting.  [If you want to use multiple "other elements" that's also fine].
  
Also: Show two visuals to explore or support what you find.

One visual can any sort of wordcloud or pplot.

The other visual must be based on the new visual types shown in the tidy text mining book: see links below
 * Section 2.4: Most positive common positive and negative words (new method of doing a bar)
 https://www.tidytextmining.com/sentiment#most-positive-negative
 
 * Section 2.5: **Comparison** wordcloud (not a regular wordcloud, the "positive/negative" one)
 https://www.tidytextmining.com/sentiment#wordclouds
 
  

#### WRITE: Provide your goal (Write your question & why it might be interesting) 
*Just write in the space below


### DO IT:
```{r}
# data prep/analysis code


```


```{r}
# viz 1


```


```{r}
# viz 2


```




# Final Thoughts
*This can be commentary only or commentary plus code/visual*  
*It does not need to be long.  Quality over quantity*

