---
title: 'Assignment #0 - Data manipulation review'
output:
  html_document:
    df_print: paged
---

# Instructions
Add code below to perform the requested actions.  
Submit your final result as a knitted HTML file to Canvas.  

Free startup
```{r startup, include = F}
library(dplyr)
library(tidyr)
glimpse(starwars)
```


## Q1
Limit the starwars dataset to only include the columns  
*name, height, mass, hair_color, eye_color, skin_color, and species*  
and assign it to the object sw  
Hint for this and following: https://r4ds.had.co.nz/transform.html
```{r Q1}

sw <- starwars #your code here

#show the output
colnames(sw)
```

## Q2
Limit your new dataset, sw, further to only include characters who are taller than 181cm
```{r Q2}
sw <- sw #your code here

#show some output
min(sw$height)
```


## Q3
Order your revised dataset alphabetically by name, and show the first 10
```{r Q3}
sw_first10 <- sw  #your code here

#show some output (adjust/add to this code if necessary)
sw_first10
```


## Q4
Create a new column for the base starwars data that converts cm to inches, called "height_inch"  
The value should be height divided by 2.54 
```{r Q4}
sw2 <- starwars #your code here

#show some results
paste("Characters who are", sw2$height[1], "in cm are", sw2$height_inch[1], "in inches")
```
## Q5
For each species as a group, calculate and show the *average height in inches* and the *average mass*.  
Make sure each answer is a number (not NA) but NaN is acceptable.
```{r Q5}
sw_metrics <- sw2  #your code here

#show some results 
sw_metrics
```

## Q6
Make the sw data more vertical (pivot longer) by combining the information 
from eye_color, hair_color, and skin_color  
into one column (called "descriptor" with the original column name)   
and a second column (called "color" with the original column value).  
Hint: https://r4ds.had.co.nz/tidy-data.html

```{r}
sw_pivoted <- sw  #your code here

#show some results
sw_pivoted
```

## Q7
Make a rough cut of name into 2 columns, one for first name and one for last name, where the split point is a space.  
If a name has multiple spaces, only split it once (at the first space)  
If a name as no space, it should not be split (and last name will end up as null)  
*Hint: if you're using separate, check documentation for the "extra" argument  
```{r Q7}
sw_names <- sw  #your code here

#show results for humans, base-R style
sw_names[sw_names$species == 'Human',]
```

## Q8  
How many characters only have 1 name?  
There are a several methods to figure this out where R can count for you  
```{r Q8}

sw_names #your code here

```

