---
title: "Regular Expressions Lab Solutions"
author: "Unstructured Data Analytics (ITAO/MSSA)"

output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
    df_print: paged
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

### Click to expand the code chunks

In this exercise only requires the `stringr` package. Install and load it as needed.

```{r echo=TRUE, message=FALSE, warning=FALSE}
#library(tidyverse)
library(stringr)
```

## Part 1: Practice with the `words` data
Contain "oo"
```{r}
str_subset(words, pattern = "oo")
```

Start with "g"
```{r}
str_subset(words, pattern = "^g")
```

End with "d"
```{r}
str_subset(words, pattern = "d$")
```

Start with a vowel
```{r}
str_subset(words, pattern = "^[aeiou]")
```


## Part 2: Practice with the `fruit` data
Ends with a vowel
```{r}
str_subset(fruit, "[aeiou]$")
```

Has two words
```{r}
#simple using character classes 
str_subset(fruit, "\\w\\s\\w")

#another way to use character classes
str_subset(fruit, "[:alpha:]+\\s[:alpha:]+")

#simpler but riskier method
#simple because it only uses one thing
#risky because it would match with extraneous beginning or end spaces, if they existed in the dataset
str_subset(fruit, "\\s")


```
Has 2 "r" in a row
```{r}
str_subset(fruit, "r{2}")
```



Are single words
```{r}
#one way (only letters)
str_subset(fruit, "^[:alpha:]+$")

#another way (only non-spaces)
#would match punctuation or numbers though
str_subset(fruit, "^[^\\s]+$")

#another way
#some number of "no spaces" at the start and no space at the end either
#could still match punctuation or numbers
str_subset(fruit, "^[^\\s]*[^\\s]$")
```

Has any doubled letter
```{r}
#ok so this was hard, because it relies on groups with backreferencing... and we didn't talk about back referencing
#but i wanted you to try and think about it
#this says look for a group of "any item" (.) 
#followed by a reference to that group  \\1 
str_subset(fruit,
            pattern = "(.)\\1")
```


Match 2 "p"s that are not consecutive
```{r}
#this is enough. pepper is allowed to match because it has "pep" in it.
str_subset(fruit,
           pattern = "p.+p")

#if you want to also EXCLUDE words with consecutive pp, that's harder
#a p, followed by an number of not-p, followed by a p, followed by either more not-p or end-of-string
str_subset("paap",
          pattern = "p[^p]+p([^p]+|$)")

```
## Part 3: Practice with the `sentences` data
Contains a ? or !
```{r}
str_subset(sentences, pattern = "\\?|\\!")
```

Contains "fence" or "over" as complete words
```{r}
#if you just look for fence|over, you'll pull in words like "cover"
str_subset(sentences, pattern = "\\Wfence\\W|\\Wover\\W")
```

not X becomes not_X
```{r}
#str_replace, as you may have noticed, gives back ALL sentences with or without changes
#so to simplify the output, it's nice to find the sentences first and save them into a smaller set

sentences_not <- str_subset(sentences, pattern = " not ")

str_replace(sentences_not, pattern = " not ", replacement = " not_")

#you can probably do this in a fancier way, but simple characters strings seem to work
```

Contains "fence" and "over"
```{r}
#to be safe, include \\W, but for this set of data it doesn't matter
str_subset(sentences, pattern = "(.*\\Wfence\\W.*\\Wover\\W.*)|(.*\\Wover\\W.*\\Wfence\\W.*)")
```


## Part Whoops: Challenge (with your own sample strings)
  
Apparently I took these out of the instructions... try them if you want!  
  
Replace "ate" with "8"  
"late", "mate", "abate", "cake"
```{r}
str_replace(c("late", "mate", "abate", "cake"), 
            pattern = "ate",
            replacement = "8")
```


Replace any doubled letter with a single letter   
"boo", "moo", "baa", "kitten", "hand"
```{r}
#again, this requires groups with back references
str_replace_all(c("boo", "moo", "baa", "kitten", "hand"),
                pattern = "(.)\\1",
                replacement = "\\1")
```



## Part 4: If you want to test yourself more (for fun!)
Basic pig latin  
"apple", "pineapple", "mango"
```{r}
#this requires MORE groups with references.  Each set of () can be called as a numbered group (in order)
str_replace(c("apple", "pineapple", "mango"),
           pattern = "^(.)(.*)$",
           replacement = "\\2\\1ay")
```
Consonant-only pig latin  
"apple", "pineapple", "mango"
```{r}
str_replace(c("apple", "pineapple", "mango"),
           pattern = "^([^aeiou])(.*)$",
           replacement = "\\2\\1ay")
```

Two-word pig latin  
"pumpkin pie", "vanilla cake", "sugar cookie"
```{r}
str_replace(c("pumpkin pie", "vanilla cake", "sugar cookie"),
           pattern = "^([^aeiou])(.*)\\s([^aeiou])(.*)$",
           replacement = "\\2\\1ay \\4\\3ay")
```


