---
title: "Sentiment Analysis (Word Level)"
author: "Unstructured Data Analytics (UDA)"

output: 
  html_document:
    toc: true
    toc_float: true
    theme: spacelab
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In this exercise, we will make use of the `tidyverse`, `tidytext` and `textdata` packages. Install and load these packages as needed.

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidytext)
library(textdata)
```

### 1. The Word-level Sentiment Lexicons
The `get_sentiments()` function from the `tidytext` package allows us to access the built-in `bing` sentiment lexicon as well as others provided through the `textdata` package.

To get the `bing` sentiment lexicon, we pass the lexicon name to the `get_sentiments()` function.

```{r}
get_sentiments("bing")
```

We can do the same for the `afinn` sentiment lexicon (Note that you will need to answer "Yes" in the console by typing "1" in order for this lexicon and others like it to be downloaded).

```{r}
get_sentiments("afinn")
```

Notice the difference in the way the data is represented between these two lexicons? Let's get a count of the tokens from all four sentiment lexicons to get a big picture view of how they compare to each other.

```{r}
get_sentiments("bing") %>%
  count(sentiment)

get_sentiments("afinn") %>%
  count(value)

get_sentiments("nrc") %>%
  count(sentiment)

get_sentiments("loughran") %>%
  count(sentiment)
```

**Question:** What differences do you notice among the four lexicons?

### 2. Word-level Sentiment Analysis

To illustrate how word-level sentiment analysis works, let's begin with a very simple example.

```{r}
statement <- "This is amazing. The weather is wonderful today"
statement
```

The first thing we need to do with our text is convert it into a tibble and then tokenize by word (this should be familiar to you).

```{r}
tokens <- tibble(text = statement) %>% 
  unnest_tokens(output = word, input = text, "words")
tokens
```

Next, we join our tokenized data with the `bing` sentiment lexicon.

```{r}
tokens %>%
  inner_join(get_sentiments("bing"), by = "word")
```

Of the 8 words in our text, only 2 of them exist in the `bing` sentiment lexicon.

Finally, we get a sum of the words by sentiment to determine if the text is mostly positive or mostly negative.

```{r}
tokens %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(sentiment)
```

As you can see, based on the `bing` sentiment lexicon, our sentence is considered positive.


Let's try another example with a few more words.

```{r}
statement <- c("I have a few thoughts about the ITAO major.", 
               "I dislike programming in general.",
               "However, I really do like R.",
               "I do kind of love machine learning.")

tokens <- tibble(text = statement) %>% 
  unnest_tokens(output = word, input = text)

tokens %>%
  inner_join(get_sentiments("bing"), by = "word") 
```

This time, let's use the `pivot_wider()` command to pivot our data so we can subtract the negative sentiment count from the positive sentiment count.

```{r}
tokens %>%
  inner_join(get_sentiments("bing"), by = "word") %>%
  count(sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)
```

The sentiment score is positive, therefore the entire text has a positive sentiment.

What if we used the `nrc` lexicon?

```{r}
tokens %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  count(sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)
```

The sentiment score is also positive with the `nrc` lexicon. 

**Question:** Why did we get different sentiment scores between the `bing` and `nrc` lexicons?

A closer look at the results show that we have one negative word and four positive words. Let's see what these words are.

```{r}
tokens %>%
  inner_join(get_sentiments("nrc"), by = "word") %>%
  filter(sentiment %in% c("positive","negative"))
```